### 2016-05-26 Paper_v2.tex
Model ID paragraph

\subsection*{Formulation of a reduced order complement model}
We developed a reduced order human complement network consisting of the most crucial steps of the human complement system (Fig. \ref{fig-schematic}). The core of our model was based upon the experimental measurements of Morad and coworker's earlier work \cite{morad2015time}, we only consider the activation of complement system through the alternate and the lectin pathways. In doing so we aim to capture a complex biological phenomenon using a few simple ordinary differential equations.  A trigger event initiates the lectin pathway in the presence of zymosan, which activates the cleavage of C2 and C4 into C2a and C2b, and C4a and C4b respectively. Classical Pathway (CP) C3 convertase (C4aC2b) is a combination of C4a and C2b, which catalyzes the cleavage of C3 into C3a and C3b. Similarly, the activation of the alternative pathways happens through the spontaneous hydrolysis of C3 which facilitates the cleavage of C3. C3b then could combine with with C3 to form alternate pathway (AP) C3 convertase. Both C3 convertases catalyze the cleavage of C3 into C3a and C3b, and C3b can then combine with either CP or AP C3 convertase to form C5 convertase, CP or AP respectively that is responsible for the cleavage of C5 to C5a and C5b. Lectin pathway activation was approximated using a combination of saturation kinetics and Hill-like function control functions. These control coefficients then modified the rates of model processes at each time step. Hill-like transfer functions $0 \le \mathbf{f (Z)}  \le 1$ quantified the contribution of components upon a target process, in this study, $\mathbf{Z}$ represents the abundance of the initiator. Taken together, while the reduced order human complement model encodes significant biological complexity, it is highly compact (consisting of only 18 differential equations). Thus, it will serve as an excellent proof of principle example to study the reduction of a highly complex human subsystem.

\subsection*{An ensemble of complement models was estimated using dynamically dimensioned search.}
A critical challenge for any dynamic model is the estimation of kinetic parameters. We estimated
kinetic and control parameters in a hierarchical fashion using  two $in~vitro$ time-series human complement data sets with and without zymosan present. The residual between simulation and experimental measurements were minimized  using dynamically dimensioned search (DDS). An initial parameter set was initialized with randomized kinetic and control parameters and allowed to search for parameter vectors that minimized the residual. Knowing that the kinetic and control parameters of the lectin pathway does not affect the dynamics of the alternate pathway, we used a hierarchal approach that estimated the parameters for the alternative pathway and lectin pathway separately. For the alternative pathway, we utilized the time-course experimental measurements of Morad and coworkers \cite{morad2015time} of C3a and C5a in the absence of zymosan and only allowed the alternative parameters to vary (Fig. \ref{fig-fit} A and B). The estimated alternate parameters was then fixed for the determination of lectin pathway parameters. The training for the lectin parameters, we used the experimental measurements of C3a and C5a in the presence of $1$ g of zymosan published by Morad et al \cite{morad2015time} (Fig. \ref{fig-fit} C and D).
The reduced human complement model captured the behavior of the alternative and lectin pathways through the time-course abundance of C3a and C5a (Fig. \ref{fig-fit}).
However we were not able to capture the curvature of the C5a alternative (Fig. \ref{fig-fit}).
The decreasing slope of the experimental measurements may be an indication of the decreasing cofactors that are required for the spontaneous hydrolysis in the alternative pathway, which we neglected.  Taken together, the model identification results suggested that our reduced order approach could reproduce a panel of lectin pathway initiation data sets in the neighborhood of physiological factor and inhibitor concentrations. However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.

>>>>

\subsection*{Formulation of the reduced order complement model.}
The reduced order complement model described the alternate and lectin pathways (Fig. \ref{fig-schematic}).
A trigger event initiated the lectin pathway, which activated the cleavage of C2 and C4 into C2a, C2b, C4a and C4b respectively.
Classical Pathway (CP) C3 convertase (C4aC2b) then catalyzed the cleavage of C3 into C3a and C3b.
On the other hand, activation of the alternative pathway was initiated through the spontaneous hydrolysis of C3 into C3a and C3b.
The C3b fragment then recombined with C3 to form the alternate pathway (AP) C3 convertase.
Both the CP and AP C3 convertases catalyzed the cleavage of C3 into C3a and C3b.
A second C3b fragment could then bind with either the CP or AP C3 convertase to form the CP (or AP) C5 convertase.
The C5 convertase catalyzed the cleavage of C5 into the C5a and C5b fragments.
Lectin pathway activation was approximated using a combination of saturation kinetics and non-linear transfer functions which modified the rates of model processes at each time step. These transfer functions allowed a significant reduction in the size of the model, while maintaining the biological performance.
Thus, while the reduced order complement model encoded significant biological complexity, it was highly compact consisting of only 18 differential equations and 28 model parameters.
Next, we estimated an ensemble of model parameters from time series measurements of the C3a and C5a complement proteins.

\subsection*{Estimating an ensemble of reduced order complement models.}
A critical challenge for any dynamic model is the estimation of model parameters.
We estimated the complement model parameters in a hierarchical fashion using two \textit{in~vitro} time-series data sets generated with and without zymosan, a lectin pathway activator \cite{morad2015time}.
The residual between model simulations and experimental measurements was minimized using the dynamic optimization with particle swarms (DOPS) approach,
starting from an initial random parameter guess.
Knowing the parameters of the lectin pathway did not affect the dynamics of the alternate pathway,
we used a hierarchal approach that estimated the alternative and lectin pathway parameters separately.
For the alternative pathway, we used the C3a and C5a measurements of Morad and coworkers \cite{morad2015time} in the absence of zymosan, and only allowed the alternative parameters to vary (Fig. \ref{fig-fit} A and B).
The putative alternate parameters were then fixed, and the lectin parameters were estimated.
Lectin parameters were estimated from C3a and C5a measurmemts in the presence of 1g zymosan (Fig. \ref{fig-fit} C and D).
The reduced complement model captured the behavior of the alternative and lectin pathways (Fig. \ref{fig-fit}).
Taken together, the reduced order approach reproduced a panel of lectin pathway initiation data sets in the neighborhood of physiological factor and inhibitor concentrations.
However, it was unclear whether the reduced order model could predict new data, without updating the model parameters.

%However, we failed to fully capture the curvature of C5a alternative measureme (Fig. \ref{fig-fit}).
%The decreasing slope of the experimental measurements may be an indication of the decreasing cofactors that are required for the spontaneous hydrolysis in the alternative pathway, which we neglected.
